---
title: "Assignment5_Problem3"
author: "Supriya Savalkar"
date: "10/27/2021"
output:
  pdf_document:
    latex_engine: xelatex
---
Suppose we collect data for a group of students in a statistics class with variables:  
X1 = hours studied,  
X2 = undergrad GPA,  
X3 = PSQI score (a sleep quality index), and  
Y = receive an A.  
We fit a logistic regression and produce estimated coefficient, b = −7, b1 = 0.1, b2 = 1, b3 = -.04.  
a. Estimate the probability  that a student who studies for 30 h, has a PSQI score 
of 11 and has an undergrad GPA of 3.0 gets an A in the class. Show your work.  
X1=30h
X2=3.0
X3=11
I have referred https://rpubs.com/ppaquay/65560 for solving the problem.
The equation of predicted probability is : p(x)= (e^(b0+b1*X1+b2*X2+ b3*X3)/( 1+e^(b0+b1*X1+b2*X2+ b3*X3))
Numerator = e^(b0+b1*X1+b2*X2+ b3*X3)
Denominator = 1+e^(b0+b1*X1+b2*X2+ b3*X3)= 1+Numerator

```{r}
Numerator <-(2.71828 ^(-7 + (0.1)*(30) + (1)*(3) + (-0.04)*(11))) ## e is Euler's number = 2.71828
Denominator <- 1 + Numerator
Probability <- Numerator/Denominator
Probability
```
The probability of student getting an A is 19.15455%

b) How  many  hours  would  the  student  in  part  (a)  need  to  study  to  have  a  60  % 
chance of getting an A in the class? Show your work. 
p(x)= (e^(-7+0.1*X1+1*3+(-0.04)*11)/( 1+e^(-7+0.1*X1+1*3+(-0.04)*11))=0.6
Solving this equation, e^(-7+0.1*X1+1*3+(-0.04)*11)=1.5,
taking log on both sides 
-7+0.1*X1+1*3+(-0.04)*11=0.405
-4.44+0.1X1=0.05
X1= (0.405+4.44)/0.1

```{r}
X1= (0.405+4.44)/0.1
X1
```
Student needs to study 48.45hours

c) How many hours would a student with a  3.0 GPA and a PSQI score of 5 need 
to study to have a 50 % chance of getting an A in the class? Show your work. 
X1=?
X2=3.0
X3=5
p(x)=0.5

p(x)= (e^(-7+0.1*X1+1*3+(-0.01)*5)/( 1+e^(-7+0.1*X1+1*3+(-0.04)*5))=0.5
solving the above equation
e^(-7+0.1*X1+1*3+(-0.04)*5)=1
taking log on both sides
-7+0.1*X1+1*3+(-0.04)*5=0
X1=(4.2)/0.1

```{r}
X1=(4.2)/0.1
X1
```

Student needs to study 42hours
